---
title: "Week 3 Notebook - 2/7/2018"
output: html_notebook
---

# Class on Data Types

```{r Types}
#inside curly brackes, r then space, then it will start acceapting arguments
#by putting {r NAME} you are naming your chunk (shows up at the bottom of this pane)

2 + 2
```

# Data Objects

Vector - Atomic (can only be one type) by default if R can't recognize what type of data you're importing, it turns data into String

```{r Vectors}
#storing data or a function
#nouns for data, verbs for functions (*cans still name variables with . ex: "vector.1", but can be confused with functions)

vector_1 <- c(1,2,3,4) #c for combine/concatenate
vector_1

```


```{r Vector2}
# This is a different, longer vector

(vector_2 <- c(1:60)) # colon : is for a sequence
# if you put parenthesis around the whole fucntion, becomes self-executing function, so when you run it will also print 

```

Strings

```{r Strings}
(string_1 <- c("a", "b", "c")) #to declare something a string, each element has to be in its own quotes
(string_2 <- c("abc","def", "ghi"))

typeof(string_1) #to verify the data type of a vector
length(string_2) # to show number of items in a vector

long_string <- c(LETTERS)
long_string
```

Logicals

```{r logicals}
#true/false vectors
#when adding true/false to integers, they are converted to 1/0
t_f <-  c(TRUE, FALSE, T, F)
t_f

```

Factors - similar to true/false- grouping factos (ex. genus name)

```{r factors}
factor_1 <-  c("a","b","a","b","c") #start with string
factor_2 <- as.factor(factor_1) #use coercion to make it a factor

```



Basic math operators

$ antyhing in between dollar signs is seen as an equation and will be printed $
$+, -, *, /$
OR use \
\+ to treat the operation as a character


Matrix Operations need % to perform matrix algebra: 
%*%
%+%

without the %, r treats matrixes as a 1:1 array, working through the grid, does not use matrix algebra

```{r add vectors}
vector_add <- vector_1 + vector_1
vector_add

vector_mult <-  vector_1 * vector_1
vector_mult

vector_add_2 <- vector_1 + vector_2
vector_add_2  # r is taking the shorter vector_1 and repeating it over and over to match up with the longer (60 element) vector_2, called "recycling" - r is always going to make the 2 vectors match
```

Arrays

```{r arrays}

array_1 <- rbind(vector_1, c(5:8))
array_1

array_2 <- cbind(vector_1, c(5:8))
array_2
```

```{r dimensions}

length(array_1) #only shows total number of elements
length(array_2)

dim(array_1) #shows rows, the columns. ex: 2 rows, 4 columns
dim(array_2)

```


```{r adding arrays}
#arrays have to be the same shape and size
```


## Coercion

All Vectors can only be **One Type**
```{r}

string_3 <- c(1,2,3,4.5,"a") #changes everything to string when mixed with a character - promoting all things to something common
typeof(string_3)

vector_4 <- c(1,2,3.3,4.8)
#same for mixed integer and float, will change everything to float/double
#if using as.integer coercion, will always round down 
```


##Indexing vectors

vectors and arrays can be indexed to find specific elements

```{r Indexing}
length(vector_4)
#find the third element of the vector_4
vector_4[3]

#find multiple elementsusing a range
vector_1[1:3]
vector_2[1:10]

#pull random elements from a vector, or pull out of order
vector_1[c(3,1)]

#use a vector to subset another vector
vector_2[vector_1]

#multi-dimensional array
array_1[vector_1] #just takes the first four working in a square
array_1

array_1[2,4]#second row, 4th element
array_1
```










